Automated Program Verification and Synthesis of Embedded \& Cyber-physical Systems: Advances and Challenges

### 2017-08-17 paper.tex
reviewed section 3.2
Signed-off-by: Lucas Cordeiro <lucasccordeiro@gmail.com>

The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$. In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$, as starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{Abate17}. As a result, a given candidate program $p$ is iteratively refined, in order to match the specification $\sigma$ more closely. Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis that are currently used in practice~\cite{David15}. Figure~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the basic architecture of the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement~\cite{Alur13}.

The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{F})$, where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers. The ground terms are interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT's bit-vectors part. The phases {\sc Synthesize} and {\sc Verify} interact via a finite set of test vectors {\sc inputs} that is incrementally updated. Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$). If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the phase {\sc verify}, in order to check whether it is a full solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$). If this is the case, then the algorithm terminates; otherwise, additional information is provided to the phase {\sc synthesize}, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again. 
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{figure3.jpg}
	\caption{CounterExample Guided Inductive Synthesis (CEGIS).}
	\label{Counter-Example-Guided-Inductive-Synthesis}
\end{figure}

One may notice that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.  Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$. In order to avoid exploring all possible values, machine learning techniques can be used in the phase {\sc Synthesize}, with the goal of learning from experience, {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.

>>>>

The basic idea of program synthesis is to automatically construct a program $P$ that satisfies a correctness specification $\sigma$. In particular, program synthesis is automatically performed by engines that use a correctness specification $\sigma$, as starting point, and then incrementally produce a sequence of candidate solutions that satisfy $\sigma$~\cite{David15,Abate17}. As a result, a given candidate program $p$ is iteratively refined, in order to match the specification $\sigma$ more closely. Counter-Example Guided Inductive Synthesis (CEGIS) represents one of the most popular approaches to program synthesis that are currently used in practice~\cite{David15}. Figure~\ref{Counter-Example-Guided-Inductive-Synthesis} shows the basic architecture of the CEGIS approach, which has close connections to algorithmic debugging using counterexamples and abstraction refinement~\cite{Alur13}.

The correctness specification $\sigma$ provided to the program synthesizer is of the form $\exists \vec{F} .  \forall \vec{x}.  \sigma(\vec{x}, \vec{F})$, where $\vec{F}$ ranges over functions, $\vec{x}$ ranges over ground terms, and $\sigma$ is a quantifier-free formula typically supported by SMT solvers. The ground terms are interpreted over some finite domain $\mathcal{D}$, where $\mathcal{D}$ can be encoded using the SMT's bit-vectors part. The phases {\sc Synthesize} and {\sc Verify} interact via a finite set of test vectors {\sc inputs} that is incrementally updated. Given the correctness specification $\sigma$, the {\sc Synthesize} procedure tries to find an existential witness $\vec{F}$ satisfying the specification $\sigma(\vec{x}, \vec{F})$, for all $\vec{x}$ in {\sc inputs} (as opposed to all $\vec{x} \in \mathcal{D}$). If {\sc synthesize} succeeds in finding a witness~$\vec{F}$, this witness is a candidate solution to the full synthesis formula, which is passed to the phase {\sc verify}, in order to check whether it is a proper solution ({\it i.e.}, $\vec{F}$ satisfies the specification $\sigma(\vec{x}, \vec{F})$ for all $\vec{x}\in\mathcal{D}$). If this is the case, then the algorithm terminates; otherwise, additional information is provided to the phase {\sc synthesize}, in the form of a new counterexample that is added to the {\sc inputs} set and the loop iterates again. 
%
\begin{figure}[h]
	\centering
	\includegraphics[width=\textwidth]{figure3.jpg}
	\caption{CounterExample Guided Inductive Synthesis (CEGIS).}
	\label{Counter-Example-Guided-Inductive-Synthesis}
\end{figure}

One may notice that each iteration of the CEGIS loop adds a new input to the finite set $\text{\sc inputs}$ that is used for synthesis.  Given that the full set of inputs $\mathcal{D}$ is finite, this means that the refinement loop can only iterate over a finite number of times; however, $\mathcal{D}$ can represent a large number of elements for the finite set $\text{\sc inputs}$. In order to avoid exploring all possible values, machine learning techniques can be used in the phase {\sc Synthesize}, with the goal of learning from experience (input-output samples), {\it i.e.}, learning from counterexamples provided by a verification oracle~\cite{Alur13}.
